---
title: "Pune Tree Census August 2019(Leaflet Assignment)"
author: "Rishikesh Pillay"
date: "6/30/2021"
output: html_document
runtime: shiny
---


# Overview 
My main motive is to understand the Tree census or record of my hometown Pune city .I am proud to say that Pune city is smart city and I got this data from website .Note - Data is updated on August 2019

- [Pune Data-Store](http://opendata.punecorporation.org/Citizen/CitizenDatasets/Index) 
- [Data File](http://opendata.punecorporation.org/Citizen/CitizenDatasets/DownloadDataset/483?filepath=%2FDocuments%2F483%2FPune%20Tree%20Census%20August%202019.zip)

This data include many features like location of trees , size canopy , common names etc.
I will used this data to hopefully plot it on interactive map using Leaflet library

# Data Reading and Cleaning 

```{r Data reading, results='hide',  warning=FALSE,  message=FALSE, cache=TRUE}

library(readr)
library(dplyr)
library(shiny)

features <-
        c(
                "id",
                "canopy_dia_m",
                "condition",
                "northing",
                "easting",
                "common_name",
                "botanical_name",
                "ward",
                "local_name",
                "economic_i",
                "road_name"
        )


df <- read_csv("Data/Pune Tree Census August 2019/p5.csv")



# Merging Data and removing raw files

df <- select(df, all_of(features))



```

Removing raw data is important or it will take unnecessary space.There are 40lakhs plus rows and 28 features. I took only features I need for ploting.

# Ward 
Due to large dataset its taking longer time and High computing power map is taking longer time that's why I decided to divide data into ward then plot them.
Wards are how here city is divide under corporate officials.

```{r Ward subseting ,cache=TRUE, warning=FALSE}

dfWard1_20 <- filter(df , ward %in% c(1:20))
dfWard21_40 <- filter(df , ward %in% c(21:40))
dfWard41_60 <- filter(df , ward %in% c(41:60))
dfWard61_77 <- filter(df , ward %in% c(61:77))
paste("Number of Rows:",nrow(df)) 
rm(df)

``` 


# Leaflet Plot


## WARD (1 to 20)

```{r leaflet_1,  warning= FALSE, cache= TRUE, eval = TRUE}

library(leaflet)

plot1 <- addTiles(leaflet())


botName <- unique(dfWard1_20$botanical_name)
for (x in botName) {
        df2 <- filter(dfWard1_20, botanical_name == x)
        
        icons <-
                icons(
                        iconUrl =  ifelse(
                                df2$condition == "Healthy",
                                "icons/banyan_icon.png",
                                ifelse(
                                        df2$condition == 'Average',
                                        "icons/average.jpg",
                                        ifelse(
                                                df2$condition == "Dead",
                                                "icons/dead2.jpg",
                                                'icons/poor.jpg'
                                        )
                                )
                        ),
                        iconWidth = 30,
                        iconHeight = 55
                )
        
        popupMesage <-
                data.frame(
                        popup = paste(
                                sep = "<br>",
                                "<b>Botanical Name:</b>" ,
                                df2$botanical_name,
                                "<b>Common Name:</b>",
                                df2$common_name ,
                                paste("<b>Canopy Diameter:</b>", df2$canopy_dia_m),
                                paste("<b>Ward:</b>", df2$ward),
                                paste("<b>Economic use:</b>", df2$economic_i) ,
                                paste("<b>Street:</b>", df2$road_name) ,
                                paste("<b>Condition:</b>", df2$condition)
                        )
                )
        
        plot1 <-
                plot1 %>% addMarkers(
                        lat = df2$northing,
                        lng = df2$easting ,
                        icon = icons,
                        clusterOptions = markerClusterOptions(),
                        popup = popupMesage$popup,
                        label = paste("Local Name:", df2$local_name, ", Ward:", df2$ward)
                )
        
        
}
rm(dfWard1_20)

#plot1

```




```{r leaflet_2, echo=FALSE, warning=FALSE, cache=TRUE}
# WARD (21 to 40)

plot2 <- addTiles(leaflet())


botName <- unique(dfWard21_40$botanical_name)
for (x in botName) {
        df2 <- filter(dfWard21_40, botanical_name == x)
        
        icons <-
                icons(
                        iconUrl =  ifelse(
                                df2$condition == "Healthy",
                                "icons/banyan_icon.png",
                                ifelse(
                                        df2$condition == 'Average',
                                        "icons/average.jpg",
                                        ifelse(
                                                df2$condition == "Dead",
                                                "icons/dead2.jpg",
                                                'icons/poor.jpg'
                                        )
                                )
                        ),
                        iconWidth = 30,
                        iconHeight = 55
                )
        
        popupMesage <-
                data.frame(
                        popup = paste(
                                sep = "<br>",
                                "<b>Botanical Name:</b>" ,
                                df2$botanical_name,
                                "<b>Common Name:</b>",
                                df2$common_name ,
                                paste("<b>Canopy Diameter:</b>", df2$canopy_dia_m),
                                paste("<b>Ward:</b>", df2$ward),
                                paste("<b>Economic use:</b>", df2$economic_i) ,
                                paste("<b>Street:</b>", df2$road_name) ,
                                paste("<b>Condition:</b>", df2$condition)
                        )
                )
        
        plot2 <-
                plot2 %>% addMarkers(
                        lat = df2$northing,
                        lng = df2$easting ,
                        icon = icons,
                        clusterOptions = markerClusterOptions(),
                        popup = popupMesage$popup,
                        label = paste("Local Name:", df2$local_name, ", Ward:", df2$ward)
                )
        
        
}

rm(dfWard21_40)

#plot2

```




```{r leaflet_3,echo=FALSE,warning=FALSE, cache=TRUE}
## WARD (41 to 60)

plot3 <- addTiles(leaflet())


botName <- unique(dfWard41_60$botanical_name)
for (x in botName) {
        df2 <- filter(dfWard41_60, botanical_name == x)
        
        icons <-
                icons(
                        iconUrl =  ifelse(
                                df2$condition == "Healthy",
                                "icons/banyan_icon.png",
                                ifelse(
                                        df2$condition == 'Average',
                                        "icons/average.jpg",
                                        ifelse(
                                                df2$condition == "Dead",
                                                "icons/dead2.jpg",
                                                'icons/poor.jpg'
                                        )
                                )
                        ),
                        iconWidth = 30,
                        iconHeight = 55
                )
        
        popupMesage <-
                data.frame(
                        popup = paste(
                                sep = "<br>",
                                "<b>Botanical Name:</b>" ,
                                df2$botanical_name,
                                "<b>Common Name:</b>",
                                df2$common_name ,
                                paste("<b>Canopy Diameter:</b>", df2$canopy_dia_m),
                                paste("<b>Ward:</b>", df2$ward),
                                paste("<b>Economic use:</b>", df2$economic_i) ,
                                paste("<b>Street:</b>", df2$road_name) ,
                                paste("<b>Condition:</b>", df2$condition)
                        )
                )
        
        plot3 <-
                plot3 %>% addMarkers(
                        lat = df2$northing,
                        lng = df2$easting ,
                        icon = icons,
                        clusterOptions = markerClusterOptions(),
                        popup = popupMesage$popup,
                        label = paste("Local Name:", df2$local_name, ", Ward:", df2$ward)
                )
        
        
}

rm(dfWard41_60)

#plot3

```




```{r leaflet_4, echo=FALSE,warning=FALSE, cache=TRUE}
## WARD (60 to 77)

plot4 <- addTiles(leaflet())


botName <- unique(dfWard61_77$botanical_name)
for (x in botName) {
        df2 <- filter(dfWard61_77, botanical_name == x)
        
        icons <-
                icons(
                        iconUrl =  ifelse(
                                df2$condition == "Healthy",
                                "icons/banyan_icon.png",
                                ifelse(
                                        df2$condition == 'Average',
                                        "icons/average.jpg",
                                        ifelse(
                                                df2$condition == "Dead",
                                                "icons/dead2.jpg",
                                                'icons/poor.jpg'
                                        )
                                )
                        ),
                        iconWidth = 30,
                        iconHeight = 55
                )
        
        popupMesage <-
                data.frame(
                        popup = paste(
                                sep = "<br>",
                                "<b>Botanical Name:</b>" ,
                                df2$botanical_name,
                                "<b>Common Name:</b>",
                                df2$common_name ,
                                paste("<b>Canopy Diameter:</b>", df2$canopy_dia_m),
                                paste("<b>Ward:</b>", df2$ward),
                                paste("<b>Economic use:</b>", df2$economic_i) ,
                                paste("<b>Street:</b>", df2$road_name) ,
                                paste("<b>Condition:</b>", df2$condition)
                        )
                )
        
        plot4 <-
                plot4 %>% addMarkers(
                        lat = df2$northing,
                        lng = df2$easting ,
                        icon = icons,
                        clusterOptions = markerClusterOptions(),
                        popup = popupMesage$popup,
                        label = paste("Local Name:", df2$local_name, ", Ward:", df2$ward)
                )
        
        
}

rm(dfWard61_77)

#plot4

```

```{r Plot}

library(shiny)
library(leaflet)

ui <- fluidPage(
        
        titlePanel("Pune Tree census 2019"),
        
      navbarPage("Tree census",
                

                                tabPanel("Ward 1-20", leafletOutput("out1")), 
                                tabPanel("Ward 21-40", leafletOutput("out2")),
                                tabPanel("Ward 41-60", leafletOutput("out3")),
                                tabPanel("Ward 61-77", leafletOutput("out4"))
                        )
)

server <- function(input, output) {
        output$out1 <- renderLeaflet(plot1)
        output$out2 <- renderLeaflet(plot2)
        output$out3 <- renderLeaflet(plot3)
        output$out4 <- renderLeaflet(plot4)
}


shinyApp(ui, server)



```

# Date
```{r Date , echo=FALSE, cache=TRUE}
Date = format(Sys.Date(), '%d %b %Y')
```
## `r Date`

## Notes

1. It may take little time to load sorry for inconvenience
2. Code for ward 21 to 77 is in Rmd file in my github [Repo]("https://github.com/Rishikesh0714/Pune-Tree-Census-August-2019")
3. I Tried to use all data but its taking too much time to render  and my R is crashing.
4. There is many more entries in files name p1, p2, p3 and p5 all 1000000 rows. This visualization is not able to render whole data .
5. Feel free to use Data From my github page.



