---
title: "Pune Tree Census August 2019(Leaflet Assignment)"
author: "Rishikesh Pillay"
date: "7/3/2021"
output: html_document
---
# Overview 
My main motive is to understand the Tree census or record of my hometown Pune city .I am proud to say that Pune city is smart city and I got this data from website .Note - Data is updated on August 2019

- [Pune Data-Store](http://opendata.punecorporation.org/Citizen/CitizenDatasets/Index) 
- [Data File](http://opendata.punecorporation.org/Citizen/CitizenDatasets/DownloadDataset/483?filepath=%2FDocuments%2F483%2FPune%20Tree%20Census%20August%202019.zip)

This data include many features like location of trees , size canopy , common names etc.
I will used this data to hopefully plot it on interactive map using Leaflet library

# Data Reading and Cleaning 

```{r Data reading, results='hide',  warning=FALSE,  message=FALSE, cache=TRUE}

library(readr)
library(dplyr)
library(shiny)

features <-
        c(
                "id",
                "canopy_dia_m",
                "condition",
                "northing",
                "easting",
                "common_name",
                "botanical_name",
                "ward",
                "local_name",
                "economic_i",
                "road_name"
        )


df <- read_csv("Data/Pune Tree Census August 2019/p5.csv")



df <- select(df, all_of(features))



```

Removing raw data is important or it will take unnecessary space.There are 40lakhs plus rows and 28 features. I took only features I need for ploting.

# Leaflet Plot



```{r leaflet_1,  warning= FALSE, cache= TRUE, eval = TRUE}

library(leaflet)

plot1 <- addTiles(leaflet()) 

icons <-
                icons(
                        iconUrl =  ifelse(
                                df$condition == "Healthy",
                                "icons/banyan_icon.png",
                                ifelse(
                                        df$condition == 'Average',
                                        "icons/average.jpg",
                                        ifelse(
                                                df$condition == "Dead",
                                                "icons/dead2.jpg",
                                                'icons/poor.jpg'
                                        )
                                )
                        ),
                        iconWidth = 30,
                        iconHeight = 55
                )
        
        popupMesage <-
                data.frame(
                        popup = paste(
                                sep = "<br>",
                                "<b>Botanical Name:</b>" ,
                                df$botanical_name,
                                "<b>Common Name:</b>",
                                df$common_name ,
                                paste("<b>Canopy Diameter:</b>", df$canopy_dia_m),
                                paste("<b>Ward:</b>", df$ward),
                                paste("<b>Economic use:</b>", df$economic_i) ,
                                paste("<b>Street:</b>", df$road_name) ,
                                paste("<b>Condition:</b>", df$condition)
                        )
                )
        
        plot1 <-
                plot1 %>% addMarkers(
                        lat = df$northing,
                        lng = df$easting ,
                        icon = icons,
                        clusterOptions = markerClusterOptions(),
                        popup = popupMesage$popup,
                        label = paste("Local Name:", df$local_name, ", Ward:", df$ward)
                )
```

# Pune Tree Census 

`r plot1`

# Date
```{r Date , echo=FALSE, cache=TRUE}
Date = format(Sys.Date(), '%d %b %Y')
```
## `r Date`

# Conclusion/Notes

1. Data is very large and I took only small subset of it for there was latency in rendering the leaflet. You can Find whole data from above source
2. My Github[Repo]("https://github.com/Rishikesh0714/Pune-Tree-Census-August-2019")
